---
title: "Relationships between food groups and eating time slots according to diabetes status in adults from the UK National Diet and Nutrition Survey (2008--2017)"
runtitle: "Food groups choice and time of consumption. "
documentclass: frontiersSCNS  # or frontiersHLTH, or frontiersFPHY
author:
  - name: Chaochen Wang
    affiliation: '1'
    etal: WANG # First author's last name. 
  - name: Suzana Almoosawi
    affiliation: '2'
  - name: Luigi Palla
    affiliation: '3,4,5'
    email: Luigi.Palla@uniroma1.it # Indicates corresponding Author
    institution: Department of Public Health and Infectious Diseases, University of Rome La Sapienza
    street: Piazzale Aldo Moro 5
    city: Rome
    # state: State XX # only USA, Australia, Canada
    zip: 00185
    country: Italy
affiliation:
  - id: '1'
    department: Department of Public Health
    institution: Aichi Medical University
    city: Nagakute
    state: Aichi # only USA, Australia, Canada
    country: Japan
  - id: '2' 
    department: Faculty of Medicine, School of Public Health
    institution: Imperial College London
    city: London
    # state: State XX # only USA, Australia, Canada
    country: UK
  - id: '3' 
    department: Department of Public Health and Infectious Diseases
    institution: University of Rome La Sapienza
    city: Rome
    # state: State XX # only USA, Australia, Canada
    country: Italy
  - id: '4' 
    department: Department of Medical Statistics
    institution: London School of Hygiene  Tropical Medicine
    city: London
    # state: State XX # only USA, Australia, Canada
    country: UK
  - id: '5' 
    department: Department of Global Health, School of Tropical Medicine and Global Health
    institution: University of Nagasaki
    city: Nagasaki
    # state: State XX # only USA, Australia, Canada
    country: Japan
output: rticles::frontiers_article
csl: frontiers.csl
bibliography: frontiers.bib
---

\begin{abstract}

Time of eating has been shown to be associated with diabetes and obesity but little is known about less healthy foods and specific time of their intake over the 24 hours of the day. In this study we aimed to identify potential relationships between foods and their eating time, and see whether these associations may vary by diabetes status. The National Diet and Nutrition Survey (NDNS) including 6802 adults (age $\geq$ 19 years old) collected 749,026 food recordings by a 4-day-diary. The contingency table cross-classifying 60 food groups with 7 pre-defined eating time slots (6-9am, 9am-12pm, 12-2pm, 2-5pm, 8-10pm, 10pm-6am) was analyzed by Correspondence Analysis (CA). CA biplots displaying the associations were generated for all adults and separately by diabetes status (self-reported, pre-diabetes, undiagnosed-diabetes, and non-diabetics) to visually explore the associations between food groups and time of eating across diabetes strata. For selected food groups, odds ratios (OR, 99\% confidence intervals, CI) were derived of consuming unhealthy foods at evening/night (8pm-6am) vs. earlier time in the day, by logistic regression models with generalized estimating equations.
The biplots suggested positive associations between evening/night and consumption of puddings, regular soft drinks, sugar confectioneries, chocolates, spirits, beers, ice cream, biscuits, and crisps for all adults in the UK. The OR (99\% CIs) of consuming these foods at evening/night were respectively 1.38 (1.03, 1.86), 1.74 (1.47, 2.06), 1.92 (1.38, 2.69), 3.19 (2.69, 3.79), 11.13 (8.37, 14.80), 7.19 (5.87, 8.82), 2.38 (1.79, 3.15), 1.91 (1.67, 2.16), 1.55 (1.27, 1.88) vs. earlier time in the day. Stratified biplots found that sweetened beverages, sugar-confectioneries appeared more strongly associated with evening/night among un-diagnosed diabetics. Foods consumed in the evening/night time tend to be highly processed, easily accessible, and rich in added sugar or saturated fat. Individuals with undiagnosed diabetes are more likely to consume unhealthy foods at night. Further longitudinal studies are required to ascertain the causal direction of the association between late-eating and diabetes status.
\tiny
 \keyFont{ \section{Keywords:} Chrononutrition, time of eating, correspondence ananlysis, NDNS RP} 

\end{abstract}


# Introduction {-}


# Results {-}

<!-- Comments should be HTML-style comments -->

## Subsection 1 {-}

You can use `R` chunks directly to plot graphs.  

```{r graph, echo = TRUE, message=FALSE, fig.show = "hide"}
x <- 0:100
set.seed(999)
y <- 2 * (x + rnorm(length(x), sd = 3) + 3)
plot(x, y)
```

## Subsection 2 {-}

Frontiers requires figures to be submitted individually, in the same order as
they are referred to in the manuscript. Figures will then be automatically
embedded at the bottom of the submitted manuscript. Kindly ensure that each
table and figure is mentioned in the text and in numerical order. Permission
must be obtained for use of copyrighted material from other sources (including
the web). Please note that it is compulsory to follow figure instructions.
Figures which are not according to the guidelines will cause substantial delay
during the production process.

# Discussion

# Disclosure/Conflict-of-Interest Statement {-}

<!--  
Frontiers follows the recommendations by the International Committee of Medical
Journal Editors (http://www.icmje.org/ethical_4conflicts.html) which require
that all financial, commercial or other relationships that might be perceived by
the academic community as representing a potential conflict of interest must be
disclosed. If no such relationship exists, authors will be asked to declare that
the research was conducted in the absence of any commercial or financial
relationships that could be construed as a potential conflict of interest. When
disclosing the potential conflict of interest, the authors need to address the
following points:

 - Did you or your institution at any time receive payment or services from a
   third party for any aspect of the submitted work?
 - Please declare financial relationships with entities that could be perceived
   to influence, or that give the appearance of potentially influencing, what
   you wrote in the submitted work.
 - Please declare patents and copyrights, whether pending, issued, licensed
   and/or receiving royalties relevant to the work.
 - Please state other relationships or activities that readers could perceive to
   have influenced, or that give the appearance of potentially influencing, what
   you wrote in the submitted work.
 -->

The authors declare that the research was conducted in the absence of any
commercial or financial relationships that could be construed as a potential
conflict of interest.

# Author Contributions {-}

<!--  

When determining authorship the following criteria should be observed:

 - Substantial contributions to the conception or design of the work; or the
   acquisition, analysis, or interpretation of data for the work; AND
 - Drafting the work or revising it critically for important intellectual
   content; AND
 - Final approval of the version to be published ; AND
 - Agreement to be accountable for all aspects of the work in ensuring that
   questions related to the accuracy or integrity of any part of the work are
   appropriately investigated and resolved.

Contributors who meet fewer than all 4 of the above criteria for authorship
should not be listed as authors, but they should be acknowledged.
(http://www.icmje.org/roles_a.html)

-->

The statement about the authors and contributors can be up to several sentences
long, describing the tasks of individual authors referred to by their initials
and should be included at the end of the manuscript before the References
section.


# Acknowledgments {-}

Funding:

# Supplemental Data 

Supplementary Material should be uploaded separately on submission, if there are
Supplementary Figures, please include the caption in the same file as the
figure. LaTeX Supplementary Material templates can be found in the Frontiers
LaTeX folder

# References

A reference list should be automatically created here. However it won't. Pandoc 
will place the list of references at the end of the document instead. There are 
no convenient solution for now to force Pandoc to do otherwise. The easiest way 
to get around this problem is to edit the LaTeX file created by Pandoc before 
compiling it again using the traditional LaTeX commands.

# Figures {-}

```{r, Figure-1, ref.label = "graph", results = "hide", echo = FALSE, message = FALSE, fig.height=4, fig.width=4, fig.align='center', fig.cap='Figure caption', out.width = "85mm", out.height = "85mm"}
# You can also refer to code chunks from above to place figures at the bottom.
```

