---
title: "Relationships between food groups, eating time slots and diabetes status in adults from the UK National Diet and Nutrition Survey (2008–2017)"
author: "Chaochen Wang^1^, Suzana Almoosawi^2^, Luigi Palla^3^"
# date: "2018年11月6日"
output: 
  redoc::redoc:
    margins: 1.0 #sets margins to 0.5 inches
    line_numbers: TRUE # adds line numbers to the document
    keep_md: TRUE
    highlight_outputs: FALSE
    reference_docx: templateAJCN.docx
bibliography: bibliography.bib
csl: the-american-journal-of-clinical-nutrition.csl
# csl: brain.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```




```{r echo=FALSE, message=FALSE}
Table_nums <- captioner::captioner(prefix = "TABLE. ")
T.ref <- function(x) {
  stringr::str_extract(Table_nums(x), "[^:]*")
}

Table.1_cap <- Table_nums(name = "Tab1", 
                        caption = "Weighted means and percentages (95% CI) of the nutrients intake according to individual level carbohydrate eating classes. 
(NDNS RP 2008/09-15/16, sample size = 6155)")

Table.2_cap <- Table_nums(name = "Tab2",
                        caption = "Weighted means, percentages (95% CIs) of the characteristics by latent class memberships in the UK adults.
(NDNS RP 2008/09-2015/16, sample size = 6155)")
# Table.3_cap <- Table_nums(name = "Tab3", 
#                         caption = "Means (standard deviations), and counts (%) of the characteristics of different types of carbohydrate eating days.")

# Figure.2_cap <- Figure_nums(name = "fig_2", 
#                         caption = "Day level latent classes solution for carbohydrate temporal eating patterns.")
# Figure.3_cap <- Figure_nums(name = "fig_3", 
#                         caption = "Multilevel latent classes solution for carbohydrate temporal eating patterns: 3 classes in day level, 3 classes in individual level.")
# Figure.4_cap <- Figure_nums(name = "fig_4", 
#                         caption = "The compositions of energy consumption within each time slot by individual level carbohydrate eating classes.")

```


1. Department of Public Health, Aichi Medical University, Nagakute, Aichi, JAPAN (C.W.);
3. Brain, Performance and Nutrition Research Centre, Northumbria University, Newcastle, UK (S.A.);
1. Department of Medical Statistics, London School of Hygiene & Tropical Medicine, London, UK (L.P.)

Corresponding Author: Luigi Palla

Present address for L.P.: Department of Medical Statistics, London School of Hygiene & Tropical Medicine, Keppel Street, London, WC1E 7HT, United Kingdom. (email: Luigi.Palla\<at\>lshtm.ac.uk)

Sources of Support: This work was supported by Grants-in-Aid for Young Scientists (grant number 19K20199 to C.W.) from the Japan Society for the Promotion of Science (JSPS).

Short running head: Time of food consumption by diabetes status.  

##### 

**Abbreviations:** 

<!-- MLCA, multilevel latent class analysis; T2D, type 2 diabetes; CVD, cardiovascular disease; NDNS RP, National Diet and Nutrition Survey Rolling Programme; PHE, Public Health England; FSA, Food Standards Agency; PSU, primary sampling units; BMI, body mass index; WC, waist circumferences; HDL, high-density lipoprotein; LDL, low-density lipoprotein; TG, triglycerides; HbA1c, hemoglobin A1c; CI, confidence interval. -->


##### 

## ABSTRACT

**Background:** Time of eating has been shown to be associated with diabetes and obesity. 

**Objective:**  To identify potential relationships between foods and eating time, and see whether these associations may vary by diabetes status.

**Design:** The National Diet and Nutrition Survey (NDNS) included 6802 adults (age $\geqslant$ 19) collected 749026 food recordings by a 4-day-diary. The contingency table cross-classifying 60 foods with the pre-defined 7-time slots were analyzed by Correspondence Analysis (CA). Biplots displaying the associations were generated for all adults and separately by diabetes status (self-reported, pre-, undiagnosed-, and non-diabetics). Odds ratios (OR, 99%CIs) were derived of consuming unhealthy foods at evening/night (20:00 - 06:00) vs. earlier in the day, by logistic regression with generalized estimating equation.

**Results:** The biplots suggested positive associations between evening/night and spirits, beer, ice cream, chocolate, sugar-confectionery, biscuits, crisps, and sweetened beverages. The OR of consuming these foods at evening/night were respectively 11.13 (8.37-14.80), 7.19 (5.87-8.82), 2.38 (1.79-3.15), 3.19 (2.69-3.79), 1.92 (1.38-2.69), 1.91 (1.67-2.16), and 1.55 (1.27-1.88) vs. earlier time. Stratified biplots found that sweetened beverages, sugar-confectionery were more likely to be consumed at event/night among undiagnosed diabetics.

**Conclusions:** Foods consumed in the evening/night time tend to be highly processed, easily accessible, and rich in added sugar or saturated fat. Individuals with undiagnosed diabetes are more likely to consume unhealthy foods at night. Further longitudinal studies are required to explore the causality.


##### 

## INTRODUCTION

The timing of energy intake has been shown to be associated with obesity and diabetes. [@almoosawi2016chrono] 

We aimed to decribe the relationship between food groups and the time of day when they were consumed, and how such relationships may vary by status of type 2 diabetes.

## METHODS 


6802 adults (2810 men and 3992 women) and 749,026 food recordings collected by the UK National Diet and Nutrition Survey Rolling Programme (NDNS RP 2008-17) were analyzed in the current study [@MRCElsieWiddowsonLaboratory2018]. The sample was randomly drawn from a list of all addresses in the UK, clustered into postcode sectors. Details of the rationale, design and methods of the survey can be found in the previous published reports [@bates2014national; @roberts2018national]. Time of the day was categorized as 7 slots: 6-9 am, 9-12 noon, 12-2 pm, 2-5 pm, 5-8 pm, and 10 pm - 6 am. Foods recorded were classified into 60 standard food groups with 1 to 10 subgroups each: the details are given in Appendix R of the NDNS offical report [@NDNSdatabase2018]. We focused on the 60 standard food groups in the current analysis. Diabetes statuses were defined as: 1) healthy if fasting glucose was lower than 6.10 (mmol/L), hemoglobin A1c (HbA1c) were less than 6.5 (%), and without self-reported diabetes or under treatment of diabetes (n = 2626); 2) pre-diabetic if fasting glucose was lower between 6.10 and 6.99 (inclusive) but without self-reported diabetes or under treatment of diabetes (n = 133); 3) undiagnosed diabetic if either fasting glucose was higher or equal to 7.00 (mmol/L) or HbA1c higher or equal to 6.5 (%) but without self-reported diabetes or under treatment of diabetes (n = 99); 4) diabetic if participant had self-reported diabetes or under treatment of diabetes (n = 227). Consequently, number of adults whose diabetes status could not be confirmed was 3717 (1519 men, 2198 women) and they were kept in the whole sample analyses. 

Correspondence analysis (CA) [@greenacre2017correspondence;@Chapman2017] was used as a tool for data mining, visualisation and hypotheses generation using half of the randomly selected NDNS diary entries data. Specifically, the contingency table generated by cross-tabulating foods and time of eating were analyzed by CA. Through CA, the 60 categories of standard food and the 7 time-slot were projected on two dimensions that could jointly contain large percentage of the $\chi^2$ deviation (or inertia) of the table. Biplots that graphically showing the association between time of day and food groups were derived for all adults and seprately according to their diabetes status. To account for the hierarchy of the data (food recorded by the same individuals who lived within the same sampling units) and to calculate population marginal odds ratios (OR), logistic regression models with generalized estimating equations (GEE) were used to test associations that were suggested from biplots generated by CA using the remaining half of the diary entries data. The marginal OR (95% confidence intervals, CI) was derived of consuming unhealthy food groups selected by CA, later in ther day (8 pm - 6 am) compared to earlier in the day. CA was conducted and biplots were generated by the following packages under R environment [@Rcoreteam]: `FactoMineR, factoextra, ` [@L__2008;@factoextra]









<!-- ### Other variables and analyses -->

<!-- Computer-assisted personal interviews were undertaken by trained interviewers to collect background information of the participants on smoking habits (current, ex-smoker, and never), ethnicity (white, non-white), education level (lower than degree/degree or above level), living with a partner or not, and other socio-demographic variables. Participants also had their height, weight, blood pressure, and waist circumferences (WC) measured by the nurses during the visit of the interview. A self-completion questionnaire - the Recent Physical Activity Questionnaire (RPAQ) [@besson2009estimating] was used to estimate physical activity. Blood samples were taken from the participants. Serum total, high-density lipoprotein (HDL), and low-density lipoprotein (LDL), cholesterol, triglycerides (TG), fasting blood glucose, hemoglobin A1c (HbA1c) were measured. HbA1c of 6.5% was used as the cut off value for diagnosing diabetes. Hypertension was defined as with systolic blood pressure of 140 mmHg or above, or diastolic blood pressure of 90 mmHg or above, or taking any medication specifically to reduce blood pressure. Body mass index (BMI), was calculated as weight in kilograms divided by height in square meters. BMI was then categorized into less than 25 kg/m^2^ (normal weight), 25 to 30 kg/m^2^ (overweight), and higher or equal to 30 kg/m^2^ (obese). -->

<!-- Individual-level point estimates and 95% confidence intervals (CIs) were estimated by applying individual, nurse visiting, and blood sample weights accordingly to adjust for differences in sample selection and the clustered survey design. Descriptive statistics of participants were presented as weighted means (95% CI) or weight percentages (95% CI). After examining the distribution of the data, the following variables were log-transformed to improve normality: fasting blood glucose, HbA1c, TC, LDL, HDL, TG and average physical activity duration per day. Weighted geometric mean (95% CI) were used for all log-transformed variables. Weighted estimates of nutrients consumption across the 7-time slots of the day were calculated for each individual-level latent class. Contributions (%) of the average energy intake within time slots were evaluated by determining the percentages of energy coming from carbohydrate, fat, protein, and alcohol intake.  -->

<!-- For continuous variables, the *F* test was used to determine differences between latent classes with Bonferroni correction to account for multiple testing across > 2 classes when applicable. For categorical variables, differences between latent classes were assessed using the adjusted Pearson $\chi^2$ test for survey data. Survey-designed multivariable logistic regression (for hypertension), linear regression models (for BMI and WC) were used to test for associations between individual-level latent classes of carbohydrate eating patterns and WC/BMI. Since diabetic participants may or may not modify their carbohydrate eating habits, we also fitted all the above-mentioned regression models restricted to those without diabetes. All statistical analyses were performed with `svyset` command as implemented in Stata software version 15.1 [@stata15].  -->

## RESULTS



## DISCUSSION 


### Conclusions


## ACKNOWLEDGEMENT 

## CONFLIT OF INTEREST (COI) STATEMENT
The authors declared no conflicts of interest.

## AUTHOR'S CONTRIBUTIONS

CW, SA, and LP: designed research and had primary responsibility for final content; CW and LP performed statistical analysis; and all authors: wrote the manuscript, read and approved the final manuscript.

#####

## REFERENCES



